---
title: "Seattle Public Library Trends of My Favorite Childhood Authors"
author: "Mark Manlimos"
date: "2/24/2023"
output: html_document
---

### Introduction

Once an integral part of my youth education, going to the library has become more optional and therefore inconvenient for my education. With the accelerated rise of technology as a medium for knowledge, libraries have become underutilized by myself an much of the youth. As a kid I loved to read the Rick Riordan's "Percy Jackson" and "The Kane Chronicles" series and J.K. Rowling's "Harry Potter" series. As I aged into late middle school, I really got into J.R.R Tolkien's "Lord of the Rings" series. 

This report looks to explore the trends of book Mediums and compare how people take in books nowadays. I also am interested in seeing how my favorite books as a kid compare to each other today and to see what medium people are choosing to absorb these books through.

I will focus on:

- Authors:
  - Specifically the three mentioned above
- Book Medium:
  - This will look different in different sections. Sometimes I will focus on specific mediums such as EBOOKS and audio books, and sometimes I will focus on just digital vs. physical
- Time:
  - The element of time will only pertain to the year 2022 as that is the limitation of my data set. 
  
My report will connect the three of these focuses and analyze information taken from combinations of the three.

### Summary Information
``````{r, echo = FALSE, warning = FALSE, message = FALSE, code = readLines("summary.R")}
```

Based on this data, I was able to conclude that in 2022, **`r largest_monthly_checkouts`**

### The Dataset

- Who collected/published the data?  
- What are the parameters of the data (dates, number of checkouts, kinds of books, etc.)?
- How was the data collected or generated?  
- Why was the data collected?  
- What, if any, ethical questions do you need to consider when working with this data?  
- What are possible limitations or problems with this data?   (at least 200 words)


### First Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The first chart that you will create and include will show the trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Show more than one, but fewer than ~10 trends
  - For example, two books, or 5 different mediums
- You must have clear x and y axis labels
- The chart needs a clear title
- You need a legend for your different line colors and a clear - legend title
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, warning = FALSE, message = FALSE, code = readLines("Chart1.R")}
```

### Second Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The second chart that you will create and include will show another trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Show more than one, but fewer than ~10 trends
  - For example, two books, or 5 different mediums
- You must have clear x and y axis labels
- The chart needs a clear title
- You need a legend for your different line colors and a clear - legend title
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, warning = FALSE, message = FALSE, code = readLines("Chart2.R")}
```

### Your Choice

The last chart is up to you. It could be a line plot, scatter plot, histogram, bar plot, stacked bar plot, and more. Here are some requirements to help guide your design:

- You must have clear x and y axis labels
- The chart needs a clear title
- You need a legend for your different line colors and a clear legend title

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, warning = FALSE, message = FALSE, fig.width = 12, code = readLines("Chart3.R")}
```


